---
title: "trigeminal-RBs"
author: "Lauren Miller, nlc"
date: "2023-07-26"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(error = TRUE, comment = NA, warnings = FALSE, errors = FALSE, messages = FALSE)
```

# Set up

## Load libraries
```{r}
library(tidyverse)
library(Seurat)
```

## Load data
```{r}
data_folder5 <- "WT30_2"

WT30_2_data <- Read10X(data.dir = data_folder5)

WT30.obj <- CreateSeuratObject(counts = WT30_2_data, project = "WT30_2", min.cells = 3, min.features = 200)
```

# 'Standard' Seurat analysis

### Assign mitochondrial percent
*NOTE (NLC):* zebrafish mitochondrial genes have a lowercase prefix, using the regex pattern '^mt-' appeared to work, however only 13 mitchondrial genes are present.
```{r}
WT30.obj[["percent.mt"]] <- PercentageFeatureSet(WT30.obj, pattern = "^mt-")

# Visualize QC metrics as a violin plot
VlnPlot(WT30.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0)
```

### Run this chunk to manually set cutoffs:
```{r, eval = TRUE}
nCount_RNA_limits <- c(1e3, 7.5e4)

nFeature_RNA_limits <- c(1900, 9000)

percent_mt_limits <- c(0, 5)
```


### Visualize QC metrics before pruning

```{r}
FeatureScatter(WT30.obj, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident") +
  geom_vline(xintercept = nCount_RNA_limits, linetype = 'dashed', linewidth = 1) +
  geom_hline(yintercept = percent_mt_limits, linetype = 'dashed', linewidth = 1)

FeatureScatter(WT30.obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident") +
  geom_vline(xintercept = nCount_RNA_limits, linetype = 'dashed', linewidth = 1) +
  geom_hline(yintercept = nFeature_RNA_limits, linetype = 'dashed', linewidth = 1)

WT30.obj_metadata <- WT30.obj@meta.data

ggplot(WT30.obj_metadata, aes(x = nCount_RNA, y = nFeature_RNA)) +
  geom_point(alpha = 0.1) +
  geom_density2d() +
  geom_vline(xintercept = nCount_RNA_limits,  linetype = 'dashed', linewidth = 1) +
  geom_hline(yintercept = nFeature_RNA_limits, linetype = 'dashed', linewidth = 1) +
  scale_x_log10()

ggplot(WT30.obj_metadata, aes(x = nCount_RNA/nFeature_RNA, y = nFeature_RNA)) +
  geom_point(alpha = 0.1) +
  geom_density2d() +
  geom_hline(yintercept = nFeature_RNA_limits, linetype = 'dashed', linewidth = 1)
```

# Subset data

### Display counts before pruning
```{r}
table(WT30.obj@meta.data$orig.ident)
```

### Prune
```{r}
WT30.obj <- subset(WT30.obj,
                   subset =
                     nFeature_RNA > min(nFeature_RNA_limits) 
                   & nFeature_RNA < max(nFeature_RNA_limits) 
                   & nCount_RNA > min(nCount_RNA_limits) 
                   & nCount_RNA < max(nCount_RNA_limits) 
                   & percent.mt > min(percent_mt_limits) 
                   & percent.mt < max(percent_mt_limits))

```


### Display counts after pruning
```{r}
table(WT30.obj@meta.data$orig.ident)
```


### Visualize QC after pruning
```{r}
VlnPlot(WT30.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0)
```


### Normalize, identify variable features and scale data
```{r}
WT30.obj <- NormalizeData(WT30.obj)

WT30.obj <- FindVariableFeatures(WT30.obj)

WT30.obj <- ScaleData(WT30.obj, features = row.names(WT30.obj), verbose = FALSE)

WT30.obj <- RunPCA(WT30.obj, npcs = 50, verbose = FALSE)
```


### Run the standard workflow for visualization and clustering
```{r}
DimPlot(WT30.obj, reduction = 'pca')

ElbowPlot(WT30.obj, ndims = 30)

WT30.obj <- FindNeighbors(WT30.obj, reduction = "pca", dims = 1:30)
WT30.obj <- FindClusters(WT30.obj, resolution = 2)
WT30.obj <- RunUMAP(WT30.obj, reduction = "pca", dims = 1:30)

DimPlot <- DimPlot(WT30.obj, reduction = 'umap', label = T) + coord_equal()

DimPlot
```



# Directed analysis
## Identify clusters

## proliferating and mature neurons
```{r}
DefaultAssay(WT30.obj) <- "RNA"
FeaturePlot(WT30.obj, 
            features = c("mki67", "pcna"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal() 

FeaturePlot(WT30.obj, 
            features = c("snap25a", "snap25b"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal() 
```
## lateral line 
```{r}
DefaultAssay(WT30.obj) <- "RNA"
FeaturePlot(WT30.obj, features = c("ret", "gfra1b"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()

FeaturePlot(WT30.obj, features = c("alcama", "trpv1"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()
```
## trigeminal
```{r}
DefaultAssay(WT30.obj) <- "RNA"
FeaturePlot(WT30.obj, features = c("isl2b", "kitb"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()

FeaturePlot(WT30.obj, features = c("ntrk2a", "ntrk3a"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()

FeaturePlot(WT30.obj, features = c("prdm14", "trpv1"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()
```
## endocrine pancreas
```{r}
DefaultAssay(WT30.obj) <- "RNA"
FeaturePlot(WT30.obj, features = c("pax6b", "nkx2.2a"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()

## pancreas alpha, beta, delta cells
FeaturePlot(WT30.obj, features = c("gcgb", "ins", "sst2", 
                                        "arxa", "pdx1", "cdx4"),
            ncol = 3,
            min.cutoff = "q10", 
            order = F) & NoAxes() & coord_equal()
```
## motor neurons
```{r}
DefaultAssay(WT30.obj) <- "RNA"
FeaturePlot(WT30.obj, features = c("mnx1", "slc18a3a"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()
FeaturePlot(WT30.obj, features = c("hoxb9a", "ret"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()

```
## retina 
```{r}
DefaultAssay(WT30.obj) <- "RNA"
FeaturePlot(WT30.obj, features = c("atoh7", "tbx2b"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()

FeaturePlot(WT30.obj, features = c("rx2", "six6b"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()

FeaturePlot(WT30.obj, features = c("vsx1", "otx2b"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()
```
## brain
```{r}
DefaultAssay(WT30.obj) <- "RNA"
FeaturePlot(WT30.obj, features = c("ccne2", "dbx1a"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()
FeaturePlot(WT30.obj, 
            features = c("kif26ab", "slc1a2b", "stmn3"),
            ncol = 3,
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()
FeaturePlot(WT30.obj, 
            features = c("lmo3", "stox2a", "gpc1a"),
            ncol = 3,
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()
FeaturePlot(WT30.obj, 
            features = c("gabrb4", "kif5c", "apc2"),
            ncol = 3,
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()
```    
## forebrain/telencephalon
```{r}
DefaultAssay(WT30.obj) <- "RNA"
FeaturePlot(WT30.obj, 
            features = c("foxp2", "tbr1b", "bcl11aa"),
            ncol = 3,
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal() 

FeaturePlot(WT30.obj, 
            features = c("bcl11ba", "lhx9", "myt1la"),
            ncol = 3,
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal() 

FeaturePlot(WT30.obj, 
            features = c("dscaml1", "islr2", "celf5a"),
            ncol = 3,
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal() 
```
## epiphysis
```{r}
DefaultAssay(WT30.obj) <- "RNA"
FeaturePlot(WT30.obj, features = c("otx5", "aanat2", "arl3l1"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            ncol = 3,
            order = F) & NoAxes() & coord_equal()
```
## olfactory placode
```{r}
DefaultAssay(WT30.obj) <- "RNA"
FeaturePlot(WT30.obj, features = c("krt18b", "krt4", "ptgs2a"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            ncol = 3,
            order = F) & NoAxes() & coord_equal()
```
## olfactory
```{r}
DefaultAssay(WT30.obj) <- "RNA"
FeaturePlot(WT30.obj, features = c("pvalb5", "cldnh"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()
```
##otic
```{r}
DefaultAssay(WT30.obj) <- "RNA"
FeaturePlot(WT30.obj, features = c("irx1b", "irx2a"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()
```
## epibranchial
```{r}
DefaultAssay(WT30.obj) <- "RNA"
FeaturePlot(WT30.obj, features = c("phox2a", "phox2bb"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()
FeaturePlot(WT30.obj, features = c("foxi1", "pax2a"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()
```
#sympathetic nervous system
```{r}
FeaturePlot(WT30.obj, features = c("th", "dbh"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()

FeaturePlot(WT30.obj, features = c("elavl3", "dbh"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()
```
## (crainial) neural crest? 
```{r}
DefaultAssay(WT30.obj) <- "RNA"
FeaturePlot(WT30.obj, 
            features = c("snai1a", "twist1a"),
            ncol = 2,
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()

```
# epidermis? basal cells??
```{r}
FeaturePlot(WT30.obj, 
            features = c("col5a3b", "cldn1", "lamb4"),
            ncol = 3,
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()

FeaturePlot(WT30.obj, 
            features = c("cldni", "pfn1", "krt97"),
            ncol = 3,
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()
```
# muscle
```{r}
FeaturePlot(WT30.obj, 
            features = c("myod1", "chrnd"),
            ncol = 2,
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F) & NoAxes() & coord_equal()
```


### Cluster subset
```{r}
DimPlot

subset_cluster_id <- c(24,31)

#Pull the subset
WT30.obj_subset <- subset(WT30.obj, subset = seurat_clusters %in% subset_cluster_id)

# stash the original cluster identities
WT30.obj_subset@meta.data$original_seurat_clusters <- WT30.obj_subset@meta.data$seurat_clusters

# perform relatively hands-free analysis of the selected clusters
WT30.obj_subset <- FindVariableFeatures(WT30.obj_subset)
WT30.obj_subset <- RunPCA(WT30.obj_subset)
ElbowPlot(WT30.obj_subset)

WT30.obj_subset <- FindNeighbors(WT30.obj_subset, dims = 1:20)
WT30.obj_subset <- RunUMAP(WT30.obj_subset, dims = 1:20)
WT30.obj_subset <- FindClusters(WT30.obj_subset, resolution = 1)

DimPlot_subset1 <- DimPlot(WT30.obj_subset, pt.size = 2, label = T, label.size = 10) &
  NoAxes() & coord_equal()

DimPlot_subset2 <- DimPlot(WT30.obj_subset, group.by = 'original_seurat_clusters', pt.size = 2, label = T, label.size = 10) &
  NoAxes() & coord_equal()

DimPlot_subset1 + DimPlot_subset2
```

# clusters
```{r}
mycols2 <- c(
  "darkorchid4",
  "orange1",
  "deepskyblue4"
)

clusters <- DimPlot(WT30.obj_subset, 
        pt.size = 4, 
        label = F, 
        label.size = 10,
        cols = mycols2) &
  NoLegend() &
    theme(text = element_text(family = "Lato"),
        plot.title = element_text(size = 60, 
                                  face = "italic"),
        axis.title = element_text(size = 30),
        axis.text = element_text(size = 25)) &
  labs(title = "",
       x = "UMAP 1",
       y = "UMAP 2") & 
  guides(colour = guide_legend(override.aes = list(size=6)))
clusters
#ggsave("clusters.png", clusters, dpi = 1200)
```


# general trigeminal/RB markers
```{r}
#fgf13a
fgf13a <- FeaturePlot(WT30.obj_subset, 
            features = c("fgf13a"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T,
            pt.size = 4) &
  theme(text = element_text(family = "Lato"),
        plot.title = element_text(size = 60, 
                                  face = "italic"),
        legend.text=element_text(size = 35,
                                 face = "plain"),
        legend.key.size = unit(2.3, 'line'),
        legend.position = c(0.84, 0.3),
        axis.title = element_text(size = 30),
        axis.text = element_text(size = 25)) &
  labs(title = "fgf13a",
       x = "UMAP 1",
       y = "UMAP 2") 
fgf13a
ggsave("fgf13a.png", fgf13a, dpi = 1200)

#isl2a
isl2a <- FeaturePlot(WT30.obj_subset, 
            features = c("isl2a"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T,
            pt.size = 4) &
  theme(text = element_text(family = "Lato"),
        plot.title = element_text(size = 60, 
                                  face = "italic"),
        legend.text=element_text(size = 35,
                                 face = "plain"),
        legend.key.size = unit(2.3, 'line'),
        legend.position = c(0.84, 0.3),
        axis.title = element_text(size = 30),
        axis.text = element_text(size = 25)) &
  labs(title = "isl2a",
       x = "UMAP 1",
       y = "UMAP 2") 
isl2a
ggsave("isl2a.png", isl2a, dpi = 1200)

```

# specific cluster markers
```{r}
#kitb
kitb <- FeaturePlot(WT30.obj_subset, 
            features = c("kitb"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T,
            pt.size = 4) &
  theme(text = element_text(family = "Lato"),
        plot.title = element_text(size = 60, 
                                  face = "italic"),
        legend.text=element_text(size = 35,
                                 face = "plain"),
        legend.key.size = unit(2.3, 'line'),
        legend.position = c(0.84, 0.3),
        axis.title = element_text(size = 30),
        axis.text = element_text(size = 25)) &
  labs(title = "kitb",
       x = "UMAP 1",
       y = "UMAP 2") 
kitb
ggsave("kitb.png", kitb, dpi = 1200)

#tmem178
tmem178 <- FeaturePlot(WT30.obj_subset, 
            features = c("tmem178"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T,
            pt.size = 4) &
  theme(text = element_text(family = "Lato"),
        plot.title = element_text(size = 60, 
                                  face = "italic"),
        legend.text=element_text(size = 35,
                                 face = "plain"),
        legend.key.size = unit(2.3, 'line'),
        legend.position = c(0.84, 0.3),
        axis.title = element_text(size = 30),
        axis.text = element_text(size = 25)) &
  labs(title = "tmem178",
       x = "UMAP 1",
       y = "UMAP 2") 
tmem178
ggsave("tmem178.png", tmem178, dpi = 1200)

#adcyap1a
adcyap1a <- FeaturePlot(WT30.obj_subset, 
            features = c("adcyap1a"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F,
            pt.size = 4) &
  theme(text = element_text(family = "Lato"),
        plot.title = element_text(size = 60, 
                                  face = "italic"),
        legend.text=element_text(size = 35,
                                 face = "plain"),
        legend.key.size = unit(2.3, 'line'),
        legend.position = c(0.84, 0.3),
        axis.title = element_text(size = 30),
        axis.text = element_text(size = 25)) &
  labs(title = "adcyap1a",
       x = "UMAP 1",
       y = "UMAP 2") 
adcyap1a
ggsave("adcyap1a.png", adcyap1a, dpi = 1200)
```

### use hoxd3a expression to classify head vs trunk cells
```{r}
#hoxd3a
hoxd3a <- FeaturePlot(WT30.obj_subset, 
            features = c("hoxd3a"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F,
            pt.size = 4) &
  theme(text = element_text(family = "Lato"),
        plot.title = element_text(size = 60, 
                                  face = "italic"),
        legend.text=element_text(size = 35,
                                 face = "plain"),
        legend.key.size = unit(2.3, 'line'),
        legend.position = c(0.8, 0.3),
        axis.title = element_text(size = 30),
        axis.text = element_text(size = 25)) &
  labs(title = "hoxd3a",
       x = "UMAP 1",
       y = "UMAP 2") 
hoxd3a
ggsave("hoxd3a.png", hoxd3a, dpi = 1200)


# hoxd3a
Idents(WT30.obj_subset, WhichCells(object = WT30.obj_subset, expression = hoxd3a > 0.35, slot = "data")) <- 'hoxd3a.pos'

Idents(WT30.obj_subset, WhichCells(object = WT30.obj_subset, expression = hoxd3a <= 0.35, slot = "data")) <- 'hoxd3a.neg'

genes_d3a <- FindMarkers(WT30.obj_subset, ident.1 = 'hoxd3a.pos', ident.2 = 'hoxd3a.neg')

WT30.obj_subset$hoxd3a.status <- Idents(WT30.obj_subset)
```


```{r}
top20_trunk <- genes_d3a %>% filter(pct.1 > 0.7) %>% filter(pct.2 < 0.6) %>% slice_head(n = 20)
top20_head <- genes_d3a %>% filter(pct.1 < 0.5) %>% filter(pct.2 > 0.6) %>% slice_head(n = 20)

markers.to.plot <- c("hoxd3a",
"hoxa9b",
"hoxb9a",
"hoxc3a",
"hmgn3",
"hs3st1l2",
"nsg2",
"prex1",
"stox1",

"calb2a",
"col4a3bpa",
"glceb",
"itpr3",
"nefma",
"six1b",
"sncga",
"syt9b")


hoxDotPlot <- DotPlot(WT30.obj_subset, 
        assay = "RNA", 
        features = markers.to.plot, 
        cols = c("blue","red"),
        dot.scale = 12, 
        group.by = "hoxd3a.status",
        scale = F) & 
  RotatedAxis() &

  theme(text = element_text(family = "Lato", 
                            color = "#4e5355"),
       axis.title.x = element_blank(),
       axis.text.x = element_text(family = "Lato",
                                  face = "italic",
                                  color = "#4e5355",
                                  size = 20),
       axis.ticks.length = unit(0.2,"cm"),
       axis.title.y = element_blank(),
       axis.text.y = element_text(family = "Lato",
                                  color = "#4e5355",
                                  size = 20,),
       legend.title = element_text(size = 20,
                                face = "plain"),
       legend.text = element_text(size = 20,
                                face = "plain"),
       legend.key.width = unit(1, 'cm'),
       legend.key.height = unit(1, 'cm'),
       legend.position = "bottom",
       legend.justification = "center") &
  guides(colour = guide_colourbar(title = "Average Expression",
                                  title.position="top", 
                                  title.hjust = 0.5),
         size = guide_legend(title = "Percent Expressed",
                             title.position="top", 
                             title.hjust = 0.5)) &
  scale_y_discrete(labels=c("1" = "Pioneers", "0" = "Followers")) &  
  geom_point(aes(size = pct.exp), 
             shape = 21, 
             stroke = 0.9, 
             color = "#4e5355", 
              fill = alpha("#FFFFFF", 0.1)) 
hoxDotPlot
ggsave("hoxDotPlot.png", hoxDotPlot, dpi = 1200)
```

# hs3st1l2 (hox+ marker) and calb2a (hox- marker)
```{r}
#hs3st1l2
hs3st1l2 <- FeaturePlot(WT30.obj_subset, 
            features = c("hs3st1l2"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F,
            pt.size = 4) &
  theme(text = element_text(family = "Lato"),
        plot.title = element_text(size = 60, 
                                  face = "italic"),
        legend.text=element_text(size = 35,
                                 face = "plain"),
        legend.key.size = unit(2.3, 'line'),
        legend.position = c(0.84, 0.3),
        axis.title = element_text(size = 30),
        axis.text = element_text(size = 25)) &
  labs(title = "hs3st1l2",
       x = "UMAP 1",
       y = "UMAP 2") 
hs3st1l2
ggsave("hs3st1l2.png", hs3st1l2, dpi = 1200)


calb2a <- FeaturePlot(WT30.obj_subset, 
            features = c("calb2a"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = F,
            pt.size = 4) &
  theme(text = element_text(family = "Lato"),
        plot.title = element_text(size = 60, 
                                  face = "italic"),
        legend.text=element_text(size = 35,
                                 face = "plain"),
        legend.key.size = unit(2.3, 'line'),
        legend.position = c(0.84, 0.3),
        axis.title = element_text(size = 30),
        axis.text = element_text(size = 25)) &
  labs(title = "calb2a",
       x = "UMAP 1",
       y = "UMAP 2") 
calb2a
ggsave("calb2a.png", calb2a, dpi = 1200)

```





